Quantifying and Modeling Global Water Accessibility including Water Typology

### 2021-01-15 docs/manuscript/manuscript.tex
Rough Draft of the Data Section.

\citet{price2019difference} highlights the temporal dynamics of drinking water access and quality in urban slums which
is home to about 1 billion people globally. \citet{price2019difference} discusses the temporal changes in drinking water
access and quality in urban slums may influence on health risks. \citet{price2019difference} believes that monitoring
temporal dynamics should be considered over three interlinked time scales:

\begin{center}
\begin{tabular}{ c c c }
 short-term & medium term & long-term \\ 
  & &  \\  
  & &     
\end{tabular}
\end{center}
\citet{price2019difference} recommends future research to explore the short-term water access and quality monitoring. Also we should learn to recognize the role of socio-cultural factors that may impact the temporal dynamics of safe water access. Our project can pinpoint additional water accessibility variables that can be used to focus to recognize the time variations as well.  
\paragraph{}
\citet{rawas2020comparing} highlights the need for different approaches of intermittent water supplies (IWS) to monitor piped water supply perpetually in Peru. This relates and informs one mode of transportation of clean and accessible water. Our research can make inferences about if the customers receive clean water elsewhere, such as water tankers, kiosks, vendors, etc. Overall, the article found that the percent of households with (IWS) was a few percentage points greater than of the reported utilities. 
\paragraph{}
\citet{deshpande2020mapping} discusses their findings of improved overall water availability and piped water accessibility for the globe (40\% to 50.3\%). \citet{deshpande2020mapping} also found that it was the lowest increase in sub-Saharan Africa, where the accessibility was mostly concentrated in urban regions. \citet{deshpande2020mapping} focuses mostly on understanding trends in diarrhea burden, where the needs are greatest for the improvement of access to safe drinking water and sanitation facilities. Our study can benefit this research by implementing the best and effective modes of transportation of clean water to improve safe drinking water and sanitation. Even with the water pipe access increasing, they can't conclude if the people had clean water as they needed right away. Overall, the access to safe drinking water and sanitation improved globally between 2000 and 2017, but dis-proportionality undermines reaching the SDG goals. 
\paragraph{}
\citet{goal} explains specific targets that have goals and indicators to track global water accessibility. The article quantifies and provides Graphs: safe drinking water, safe sanitation, and hygiene, safe sanitation and hygiene, ambient water quality, water use efficiency, levels of freshwater stress, integrated water management, trans-boundary water cooperation, protect and restore water-related ecosystems, water and sanitation support, local participation in sanitation management. They have many ambitious plans to achieve by 2020 and 2030, and our research can find a correlation to why achieving this may be possible.
% citep (for citations fully enclosed in parentheses)
% cite (for citations without any parentheses)
% citet (for citations where the year only is in parentheses)
%%\textit{Discuss the who/what/when/where? (You already have this... just need to reorganize or improve structure)}
%Another example is \citet{fuller2016tracking}.
%\citet{rawas2020comparing} highlights the need for different approaches of intermittent water supplies (IWS) to monitor piped water supply perpetually in Peru.
%\citet{cassivi2018access} discuss time access...
%\hl{Perhaps insert a table summarizing the various dimensions of water accessibility that have been studied}
%Other examples: \citep{amit2019measuring,ho2014challenge}
\subsection{Gaps and areas for improvement}
We are interested in analyzing the impact of transportation modes (usage/ownership) on water accessibility.

@@

\section*{Data}
\hl{In this section, describe the data sources and the variables, along with data preparation procedures, variable
  selection criteria, etc. You should have a table summarizing the water accessibility variables and their
  definitions. Another table can summarize the explanatory variables (e.g. transportation-related, income-related,
  etc). Also discuss the number of countries, etc. I've started a table as an example. You can add other columns if needed.}

\begin{table}[h!]
  \centering
  \begin{tabular}{c c m{7cm}}\toprule
    \bf Variable & \bf Description & \bf Definition \\\midrule
    \texttt{ptap} & Households using a public tap/standpipe
 & Percentage of households whose main source of drinking water is a public tap/standpipe.
 \\\hline
 \texttt{bore} & Households using a tube well/borehole & Percentage of households whose main source of drinking water is a tube-well/borehole.
\\\hline
\texttt{pspr} & Households using a protected spring
& Percentage of households whose main source of drinking water is a protected spring.
\\\hline
\texttt{uspr} & Households using an unprotected spring
& Percentage of households whose main source of drinking water is an unprotected spring.
\\\hline
\texttt{pwel} & Households using a protected well
& Percentage of households whose main source of drinking water is a protected well.
\\\hline
\texttt{uwel} & Households using an unprotected well water
&Percentage of households whose main source of drinking water is an unprotected well.
\\\hline
\texttt{surw} & Households using surface water
& Percentage of households whose main source of drinking water is surface water.
\\\hline
\texttt{tles} & Households with water 30 minutes or less away round trip
&Percentage of households with water 30 minutes or less away round trip.
\\\hline 
\texttt{tmor} & Households with water more than 30 minutes away round trip & Percentage of households with water more than 30 minutes away round trip.
\\\hline
\texttt{rain} & Households using rainwater
& Percentage of households whose main source of drinking water is rainwater.
\\\hline
\texttt{phom} & Households using water piped into dwelling
&Percentage of households whose main source of drinking water is water piped into the dwelling.
\\\hline

@@

\paragraph{Preamble}
We analyzed data from the Demographic Health Survey (DHS) and the Multiple Indicator Cluster Survey (MICS) (1990-2020) to highlight water accessibility predictors in 78 countries throughout the world.

@@

>>>>

\citet{price2019difference} highlights the temporal dynamics of drinking water access and quality in urban slums which
is home to about 1 billion people globally. \citet{price2019difference} discusses the temporal changes in drinking water
access and quality in urban slums may influence on health risks. \citet{price2019difference} believes that monitoring
temporal dynamics should be considered over three interlinked time scales:
\begin{center}
\begin{tabular}{ c c c }
 short-term & medium term & long-term \\ 
  & &  \\  
  & &     
\end{tabular}
\end{center}

\citet{price2019difference} recommends future research to explore the short-term water access and quality monitoring. Also we should learn to recognize the role of socio-cultural factors that may impact the temporal dynamics of safe water access. Our project can pinpoint additional water accessibility variables that can be used to focus to recognize the time variations as well.

\citet{rawas2020comparing} highlights the need for different approaches of intermittent water supplies (IWS) to monitor piped water supply perpetually in Peru. This relates and informs one mode of transportation of clean and accessible water. Our research can make inferences about if the customers receive clean water elsewhere, such as water tankers, kiosks, vendors, etc. Overall, the article found that the percent of households with (IWS) was a few percentage points greater than of the reported utilities.

\citet{deshpande2020mapping} discusses their findings of improved overall water availability and piped water accessibility for the globe (40\% to 50.3\%). \citet{deshpande2020mapping} also found that it was the lowest increase in sub-Saharan Africa, where the accessibility was mostly concentrated in urban regions. \citet{deshpande2020mapping} focuses mostly on understanding trends in diarrhea burden, where the needs are greatest for the improvement of access to safe drinking water and sanitation facilities. Our study can benefit this research by implementing the best and effective modes of transportation of clean water to improve safe drinking water and sanitation. Even with the water pipe access increasing, they can't conclude if the people had clean water as they needed right away. Overall, the access to safe drinking water and sanitation improved globally between 2000 and 2017, but dis-proportionality undermines reaching the Sustainable Development Goals.

\citet{goal} explains specific targets that have goals and indicators to track global water accessibility. The article quantifies and provides Graphs: safe drinking water, safe sanitation, and hygiene, safe sanitation and hygiene, ambient water quality, water use efficiency, levels of freshwater stress, integrated water management, trans-boundary water cooperation, protect and restore water-related ecosystems, water and sanitation support, local participation in sanitation management. They have many ambitious plans to achieve by 2020 and 2030, and our research can find a correlation to why achieving this may be possible.
% citep (for citations fully enclosed in parentheses)
% cite (for citations without any parentheses)
% citet (for citations where the year only is in parentheses)
%%\textit{Discuss the who/what/when/where? (You already have this... just need to reorganize or improve structure)}
%Another example is \citet{fuller2016tracking}.
%\citet{rawas2020comparing} highlights the need for different approaches of intermittent water supplies (IWS) to monitor piped water supply perpetually in Peru.
%\citet{cassivi2018access} discuss time access...
%\hl{Perhaps insert a table summarizing the various dimensions of water accessibility that have been studied}
%Other examples: \citep{amit2019measuring,ho2014challenge}
\subsection*{Gaps and areas for improvement}
We are interested in analyzing the impact of transportation modes (usage/ownership) on water accessibility. We also desire to see an implementation of this statistical research piloted in one of the countries to see if the water accessibility variables predictors can improve the water accessibility for its country.

@@

\section*{Data}
\hl{In this section, describe the data sources and the variables, along with data preparation procedures, variable
  selection criteria, etc. You should have a table summarizing the water accessibility variables and their
  definitions. Another table can summarize the explanatory variables (e.g. transportation-related, income-related,
  etc). Also discuss the number of countries, etc. I've started a table as an example. You can add other columns if needed.}
\subsection*{Data Preparation}
Data structure this study were collected from the Demographic Health Survey (DHS) program. We used the STATCompiler from the DHS Program to develop customized table based on focusing on water, transportation, and wealth indicators across 78 countries and from 1990 to 2019.

We prepossess the data by differentiating the different surveys collected by the DHS program between DHS and Malaria Indicators Survey (MIS) through excel. After, we used R to develop and simplify the table with the most recent surveys for the most up to date analysis for the 78 countries. There was two sets of data of population and households. Therefore, we decided to remove the population data from the table to simplify because population and households data were identical.

Next, there were many missing values from the data. Therefore, we had to do a simple missing value analysis. We determined to avoid using water, transportation, or wealth variables which had greater than 50\% missing values. Ultimately, simplified data frame was 76 countries by 28 variables. After, two different tables are generated where one focused on the water accessibility variables and a explanatory variables. We selected 11 water accessibility variables based on the sources available such as location, time, and methodology \citep{price2019difference}. The water accessibility variable was organized
in the following table to allow for quick visualization for table 1. The other variables were used as an explanatory variables for determining the water accessibility predictors in Table 2. 
\begin{table}[h!]
  \centering
  \begin{tabular}{c c m{7cm}}\toprule
    \bf Variable & \bf Description & \bf Definition \\\midrule
    \texttt{ptap} & Households using a public tap/standpipe
 & Percentage of households whose main source of drinking water is a public tap/standpipe.
 \\\hline
 \texttt{bore} & Households using a tube well/borehole & Percentage of households whose main source of drinking water is a tube-well/borehole.
\\\hline
\texttt{pspr} & Households using a protected spring
& Percentage of households whose main source of drinking water is a protected spring.
\\\hline
\texttt{uspr} & Households using an unprotected spring
& Percentage of households whose main source of drinking water is an unprotected spring.
\\\hline
\texttt{pwel} & Households using a protected well
& Percentage of households whose main source of drinking water is a protected well.
\\\hline
\texttt{uwel} & Households using an unprotected well water
&Percentage of households whose main source of drinking water is an unprotected well.
\\\hline
\texttt{surw} & Households using surface water
& Percentage of households whose main source of drinking water is surface water.
\\\hline
\texttt{tles} & Households with water 30 minutes or less away round trip
&Percentage of households with water 30 minutes or less away round trip.
\\\hline 
\texttt{tmor} & Households with water more than 30 minutes away round trip & Percentage of households with water more than 30 minutes away round trip.
\\\hline
\texttt{rain} & Households using rainwater
& Percentage of households whose main source of drinking water is rainwater.
\\\hline
\texttt{phom} & Households using water piped into dwelling
&Percentage of households whose main source of drinking water is water piped into the dwelling.
\\\hline

@@

\subsection*{Preamble}
We analyzed data from the Demographic Health Survey (DHS) and the Multiple Indicator Cluster Survey (MICS) (1990-2020) to highlight water accessibility predictors in 78 countries throughout the world.

@@

\section*{Limitation}
%he problem of missing data is relatively common in almost all research and can have a significant effect on the conclusions that can be drawn from the data.
